Cluster Difference Imaging Photometric Survey (CDIPS) II. TOI$\,$837: A Validated Saturn-Sized Planet in IC~2602

### 2020-07-08 paper/ms.tex
update paper with IC2602 stellar params
With credit to JH for the idea to do it this way!

Co-authored-by: Joel Hartman <jhartman@astro.princeton.edu>

\title{
  Cluster Difference Imaging Photometric Survey (CDIPS) II.
  TOI$\,$837: A Validated Jupiter-Sized Planet in IC~2602
}

@@

\begin{abstract}
  We report the discovery of TOI 837.01 and its validation as a
  transiting planet.  We characterize the system using data from the
  NASA TESS mission, the ESA Gaia mission, ground-based photometry,
  and spectroscopy from CHIRON, FEROS, and Veloce.  We find that
  TOI$\,$837 is a $T=9.9$ G0 dwarf in the open cluster IC 2602.  The
  star and validated planet are therefore between 30 and 46 million
  years old.  The planet is warm ($P = 8.32\,{\rm d}$) and roughly
  Jupiter-sized.  Its transits are grazing: $b > 0.92$ at
  3$\sigma$.  From TESS photometry alone, the planetary size lies
  within 0.68--8.24 $R_{\rm Jup}$ (3$^{\rm rd}$--97$^{\rm th}$
  percentile), due to the degeneracy between the planetary size and
  the impact parameter of the transit.  From radial velocity
  monitoring, we limit the minimum mass to less than 1.20 $M_{\rm
  Jup}$ (3$\sigma$).  Grazing transits are a cause for concern, as they
  are usually indicative of astrophysical false positive scenarios.
  Our follow-up data show that such scenarios are highly unlikely.
  Our combined multi-color photometry, high-resolution imaging, and
  radial velocities rule out hierarchical eclipsing binary scenarios.
  Background eclipsing binary scenarios, though limited by speckle
  imaging, formally remain a 0.2\% possibility.  TOI 837.01 is
  therefore a validated adolescent exoplanet.  Further observations of
  its stellar obliquity, atmosphere, and mass should confirm its
  planetary nature beyond doubt, and could improve our understanding
  of how the physical and orbital properties of exoplanets change in
  time.
\end{abstract}

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.45\textwidth]{f7.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ 
  {\bf Hertzsprung-Russell diagram of \tn\ and members of \cn.}
  Members (black circles) were identified by
  \citet{cantatgaudin_gaia_2018}.  Gray circles are non-member stars
  within 5 standard deviations of the mean \cn\ right ascension,
  declination, and parallax.  $G$ denotes Gaia broadband magnitudes,
  $Bp$ Gaia blue, and $Rp$ Gaia red. 
  \label{fig:hr}
	}
\end{figure}

@@

\subsubsection{HR Diagram}

Figure~\ref{fig:hr} shows a Hertzsprung-Russell diagram of \tn, the
\cn\ cluster, and the ``neighborhood'' of spatially nearby stars.
Stars labelled as ``cluster members'' are those reported by
\citet{cantatgaudin_gaia_2018} based on Gaia DR2 positions, proper
motions, and parallaxes.  We included candidate members with formal
membership probability exceeding 10\%.  Save for a few low-mass
outliers, most members appear to be young and coeval.

\tn\ is in its expected position relative to the other members along
the cluster isochrone.  This photometrically limits the presence of
binary companions in the \tn\ system, to within perhaps half the
brightness ($\approx0.75$ magnitudes) of the target star.

Figure~\ref{fig:hr} also suggests that the membership census of \cn\
is incomplete.  We defined the reference ``neighborhood'' as the group
of at most $10^4$ randomly selected non-member stars within 5 standard
deviations of the mean \cn\ right ascension, declination, and
parallax.  We queried Gaia DR2 for these stars using
\texttt{astroquery} \citep{astroquery_2018}.  Many low-mass stars
appear above the main sequence, even though they were not identified
as 5-dimensional kinematic members through the unsupervised
\citet{cantatgaudin_gaia_2018} membership assignment process.

@@

Select properties of \tn\ from the literature and our analysis are
presented in Table~\ref{tab:starparams}.  The spectroscopic parameters
were measured from each of the six CHIRON spectra
(Section~\ref{subsec:chiron}). We then averaged the results, and
adopted the standard deviation across the spectra as the uncertainty.

We then calculated the stellar radius and reddening following
\citet{stassun_accurate_2017}.  We first derived the bolometric flux
by combining available broadband magnitudes from Gaia, Tycho-2, APASS,
2MASS, and WISE.  We then fitted the SED with the
\citet{kurucz_atlas12_2013} stellar atmopshere models, and summed to
find $F_{\rm bol}$.  When fitting the atmosphere model, we varied the
extinction ($A_{\rm V}$) and the overall normalization.  This
procedure yielded $A_{\rm V} = 0.20\pm0.03$, which independently
agrees with the average from the \cn\ isochrone fits of
\citet{randich_gaiaeso_2018}.
We then combined the spectroscopic effective temperature, bolometric
flux, and Gaia distance to determine the stellar radius, using the
Stefan-Boltzmann law.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f10.png}
	\end{center}
	\vspace{-0.5cm}
	\caption{ 
    {\bf Posterior probabilities of impact parameter, planet-to-star
    size ratio, and stellar density.} The photometric data fitted in
    Section~\ref{subsec:planet} allow for nonphysical planet sizes
    ($R_{\rm p}>2R_{\rm Jup}$; $\log(R_{\rm p}/R_\star) \gtrsim -1.6$)
    due to a classic degeneracy between the planet size and impact
    parameter.  Non-planetary sizes would imply non-planetary masses,
    which are ruled out by our radial velocity measurements.  This
    plot was made using \texttt{corner} \citep{corner_2016}.
		\label{fig:subsetcorner}
	}
\end{figure}

@@

Our joint model for the TESS and ground-based data was a
\citet{exoplanet:agol19} transit with physical and orbital parameters
shared across all transit windows, plus a local quadratic trend
allowed within each window.  The parameters and priors
are listed in Table~\ref{tab:posterior}.  In brief, we fitted for the
shared stellar parameters ($\log g, R_\star, u_0, u_1$), and the
shared planetary parameters ($t_0, P, b, \log(R_{\rm p}/R_\star)$).
There were also three free trend parameters for each of the nine
transits to account for the rotational variability.  This yielded a
total of 35 free parameters, of which 8 have physical relevance, and
27 are nuisance parameters.

@@

The posteriors from this procedure are given in
Table~\ref{tab:posterior}.  Crucially important is the point that the
transit is grazing.  The grazing geometry hinders our ability to infer
the stellar density and planet-to-star size ratio. The relevant
posterior probabilities are shown in Figure~\ref{fig:subsetcorner}.

The available data for \tn\ put us in an odd position. The transit
data provide a minimum size due to the grazing geometry; the radial
velocity data provide a maximum mass ($\approx$1.20 $M_{\rm jup}$).
We could combine the two with an empirically-calibrated planetary
mass-radius relation to force a more restrictive upper bound on the
planetary size \citep[{\it e.g.},][]{chen_probabilistic_2016}.  The
result of such an exercise would tautologically be that the planet
would be smaller than the largest known Jovian planets ($\approx
2R_{\rm Jup}$).  This would also imply that the impact parameter is
slightly below unity.  The planet's size, in our eyes, would remain
poorly constrained even if we were to do this exercise, so we opt to 
fit the photometry and radial velocities separately, to ensure clarity
about which datasets are providing which constraints.

@@

The maximum amplitude of the Rossiter-McLaughlin anomaly is
\citep{gaudi_prospects_2007}
\begin{align}
  \Delta V_{\rm RM} &\approx f_{\rm LD} \cdot \delta \cdot v\sin i \cdot \sqrt{1-b^2}
  \approx 14\,{\rm m\,s}^{-1},
\end{align}
for
\begin{align}
  f_{\rm LD} &= 1 - c_1 ( 1 - \mu ) - c_2 (1 - \mu )^2,
\end{align}
where $\mu \approx (1 - b^2)^{1/2}$, and for calculation purposes we
assumed $b=0.95$ and used stellar and transit parameters from
Tables~\ref{tab:starparams} and~\ref{tab:posterior}.  Although
challenging, for a 1.9$\,$hr transit of a $V=10.6$ star, a detection
could be achieved with modern spectrographs.  The next viable total
transit windows from Chile occur in January and February of 2021;
there are also a few visible per season from other southern locations.

>>>>

\title{
  Cluster Difference Imaging Photometric Survey (CDIPS) II.
  TOI$\,$837: A Validated Saturn-Sized Planet in IC~2602
}

@@

\begin{abstract}
  We report the discovery of TOI 837.01 and its validation as a
  transiting planet.  We characterize the system using data from the
  NASA TESS mission, the ESA Gaia mission, ground-based photometry,
  and spectroscopy from CHIRON, FEROS, and Veloce.  We find that
  TOI$\,$837 is a $T=9.9$ G0 dwarf in the open cluster IC 2602.  The
  star and validated planet are therefore between 30 and 46 million
  years old.  The planet is warm ($P = 8.3\,{\rm d}$) and roughly
  Jupiter-sized.  Combining the photometry with a prior on the stellar
  parameters derived from the cluster color-magnitude diagram, we find
  that the transits are grazing with $b=0.951\pm0.012$ and $R_{\rm p}
  = 0.77 \pm 0.07 \,R_{\rm Jup}$.  From radial velocity monitoring, we
  limit the minimum mass to less than 1.20 $M_{\rm Jup}$ (3$\sigma$).
  Grazing transits are a cause for concern, as they are usually
  indicative of astrophysical false positive scenarios.  Our follow-up
  data show that such scenarios are highly unlikely.  Our combined
  multi-color photometry, high-resolution imaging, and radial
  velocities rule out hierarchical eclipsing binary scenarios.
  Background eclipsing binary scenarios, though limited by speckle
  imaging, formally remain a 0.2\% possibility.  TOI 837.01 is
  therefore a validated adolescent exoplanet.  Further observations of
  its stellar obliquity, atmosphere, and mass should confirm its
  planetary nature beyond doubt, and could improve our understanding
  of how the physical and orbital properties of exoplanets change in
  time.
\end{abstract}

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.75\textwidth]{f7.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ 
  {\bf Hertzsprung-Russell diagram of \tn\ and members of \cn.}
  Members (black circles) were identified by
  \citet{cantatgaudin_gaia_2018}.  Gray circles are non-member stars
  within 5 standard deviations of the mean \cn\ right ascension,
  declination, and parallax.  $G$ denotes Gaia broadband magnitudes,
  $Bp$ Gaia blue, and $Rp$ Gaia red.  MIST isochrones (colored lines)
  fit the upper main sequence well, but diverge from the data for
  $M_\star \lesssim 0.7 M_\odot$. This is a known problem with PMS
  models (see Section~\ref{subsec:hr}).
  \label{fig:hr}
	}
\end{figure*}

@@

\subsubsection{HR Diagram}
\label{subsec:hr}

Figure~\ref{fig:hr} shows a Hertzsprung-Russell diagram of \tn, the
\cn\ cluster, and the ``neighborhood'' of spatially nearby stars.
Stars labelled as ``cluster members'' are those reported by
\citet{cantatgaudin_gaia_2018} based on Gaia DR2 positions, proper
motions, and parallaxes.  We included candidate members with formal
membership probability exceeding 10\%.  Most members appear to be
young and coeval.
\tn\ is in its expected position relative to the other members along
the cluster isochrone.  This photometrically limits the presence of
binary companions in the \tn\ system, to within perhaps half the
brightness ($\approx0.75$ magnitudes) of the target star.

Figure~\ref{fig:hr} suggests that the membership census of \cn\
is incomplete.  We defined the reference ``neighborhood'' as the group
of at most $10^4$ randomly selected non-member stars within 5 standard
deviations of the mean \cn\ right ascension, declination, and
parallax.  We queried Gaia DR2 for these stars using
\texttt{astroquery} \citep{astroquery_2018}.  Many low-mass stars
appear above the main sequence, even though they were not identified
as 5-dimensional kinematic members through the unsupervised
\citet{cantatgaudin_gaia_2018} membership assignment process.

Figure~\ref{fig:hr} also compares the data to the MIST isochrones
\citep{choi_mesa_2016}.  We used the
web-interace\footnote{\url{http://waps.cfa.harvard.edu/MIST/interp_isos.html},
\texttt{2020-07-08}} to interpolate isochrones at 10, 20, 30, and 40
million years. We assumed solar metallicity, and a fixed extinction
value of $A_V = 0.217$ \citep{randich_gaiaeso_2018}.  The 30 and 40
Myr models align well with the data for stars with masses ranging from
roughly $0.7$--$7\,M_\odot$.  The PMS K and M dwarf models are
bluer than observed in the Gaia photometry.  This discrepancy was noted and discussed at
length by \citet{choi_mesa_2016}.  The most plausible explanation
seems to be that strong magnetic fields in low-mass pre-main-sequence
stars inhibit convection and produce a high filling factor of
starspots \citep[{\it
e.g.},][]{stauffer_why_2003,feiden_magnetic_2013}. To conserve flux,
the stellar radius expands, and the star becomes cooler than a
1-dimensional model would predict.

@@

Select properties of \tn\ from the literature and our analysis are
presented in Table~\ref{tab:starparams}.  We calculated the stellar
parameters using two different approaches.

In ``Method 1'', we measured spectroscopic parameters from each of the
CHIRON spectra (Section~\ref{subsec:chiron}).
We then calculated the stellar radius and reddening
following \citet{stassun_accurate_2017}.  We first derived the
bolometric flux by combining available broadband magnitudes from Gaia,
Tycho-2, APASS, 2MASS, and WISE.  We then fitted the SED with the
\citet{kurucz_atlas12_2013} stellar atmopshere models, and summed to
find $F_{\rm bol}$.  When fitting the atmosphere model, we varied the
extinction ($A_{\rm V}$) and the overall normalization.  This
procedure yielded $A_{\rm V} = 0.20\pm0.03$, which agrees with the
average from the \cn\ isochrone fits of \citet{randich_gaiaeso_2018}.
Combining the spectroscopic effective temperature, bolometric flux,
and Gaia distance, we determined the stellar radius using the
Stefan-Boltzmann law.  Combining this radius the spectroscopic $\log
g$ also yields a stellar mass.

In ``Method 2'', we simply used the observed location of \tn\ in the
HR diagram and interpolated against the 40$\,$Myr isochrone.  This
method leverages the relative location of \tn\ within the narrow \cn\
isochrone to derive precise, theoretically self-consistent constraints
on all of the stellar parameters.  Although this approach would fail
for a low-mass star, \tn\ is above the stellar masses where the Gaia
photometry and isochrone models begin to diverge.

Method 1 yielded a stellar mass that seemed to be high relative to the
observed CHIRON effective temperature ($1.21 M_\odot$ to $5946\,{\rm
K}$, with relative uncertainties of a few percent on each). To avoid
poorly understood systematics,  we adopted the stellar parameters
from Method 2, and report them in Table~\ref{tab:starparams}.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f10.png}
	\end{center}
	\vspace{-0.5cm}
	\caption{ 
    {\bf Posterior probabilities of impact parameter, planet-to-star
    size ratio, and stellar density.} The 3$^{\rm rd}$ and 97$^{\rm
    th}$ percentiles in the planet-to-star size ratio correspond to
    planet sizes $0.65\,R_{\rm Jup}$ and $0.91\,R_{\rm Jup}$.  This
    plot was made using \texttt{corner} \citep{corner_2016}.
		\label{fig:subsetcorner}
	}
\end{figure}

@@

Our joint model for the TESS and ground-based data was an
\citet{exoplanet:agol19} transit with physical and orbital parameters
shared across all transit windows, plus a local quadratic trend
allowed within each window.  The parameters and priors
are listed in Table~\ref{tab:posterior}.  In brief, we fitted for the
shared stellar parameters ($\log g, R_\star, u_0, u_1$), and the
shared planetary parameters ($t_0, P, b, \log(R_{\rm p}/R_\star)$).
There were also three free trend parameters for each of the nine
transits to account for the rotational variability.  This yielded a
total of 35 free parameters, of which 8 have physical relevance, and
27 are nuisance parameters.

@@

The posteriors from this procedure are given in
Table~\ref{tab:posterior}.  The transit is grazing.  In the absence of
a strong prior on the stellar density, the grazing geometry would
hinder our ability to infer the planet-to-star size ratio.  However,
our prior on the stellar parameters from the cluster isochrone fit is
strong enough that we can derive two-sided limits on the
planet-to-star size ratio.  The relevant posterior probabilities are
shown in Figure~\ref{fig:subsetcorner}.

% The available data for \tn\ put us in an odd position. The transit
% data provide a minimum size due to the grazing geometry; the radial
% velocity data provide a maximum mass ($\approx$1.20 $M_{\rm jup}$).
% We could combine the two with an empirically-calibrated planetary
% mass-radius relation to force a more restrictive upper bound on the
% planetary size \citep[{\it e.g.},][]{chen_probabilistic_2016}.  The
% result of such an exercise would tautologically be that the planet
% would be smaller than the largest known Jovian planets ($\approx
% 2R_{\rm Jup}$).  This would also imply that the impact parameter is
% slightly below unity.  The planet's size, in our eyes, would remain
% poorly constrained even if we were to do this exercise, so we opt to 
% fit the photometry and radial velocities separately, to ensure clarity
% about which datasets are providing which constraints.

@@

The maximum amplitude of the Rossiter-McLaughlin anomaly is
\citep{gaudi_prospects_2007}
\begin{align}
  \Delta V_{\rm RM} &\approx f_{\rm LD} \cdot \delta \cdot v\sin i \cdot \sqrt{1-b^2}
  \approx 14\,{\rm m\,s}^{-1},
\end{align}
for
\begin{align}
  f_{\rm LD} &= 1 - c_1 ( 1 - \mu ) - c_2 (1 - \mu )^2,
\end{align}
where $\mu \approx (1 - b^2)^{1/2}$, and for calculation purposes we
assumed $b=0.95$ and used stellar and transit parameters from
Tables~\ref{tab:starparams} and~\ref{tab:posterior}.  Although
challenging, for a 1.84$\,$hr transit of a $V=10.6$ star, a detection
could be achieved with modern spectrographs.  The next viable total
transit windows from Chile occur in January and February of 2021;
there are also a few visible per season from other southern locations.
